---
title: "Quick Report on Gun Murders"
author: "David Pershall"
date: "11/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a quick report on 2010 gun murder rates obtained from FBI reports. The original data was obtained from [this Wikipedia page](https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state).

I am using the following libraries: tidyverse, ggplot, ggrepel, ggthemes, & dplyr.

```{r loading-libs, message=FALSE, echo=FALSE}
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(dplyr)
library(ggplot2)
```

I load the data that has already been wrangled in rda/murders.rda

```{r, echo=FALSE}
load("rda/murders.rda")
```

## Murder rate by state per 100,000 in population.

Note the large state to state variability by generating a barplot showing the gun murder rate by state:

```{r murder-rate-by-state, echo=FALSE}
murders %>% mutate(abb = reorder(abb, rate)) %>% 
  ggplot(aes(abb, rate)) +
  geom_bar(width = 0.5, stat = "identity", color = "black") +
  coord_flip()
```

## US Gun Murders 2010 average rate per million.

We compute the average murder rate per million in population. and graph all the pertinent information.

```{r average-murder-rate-per-million}
r <- murders %>% 
  summarize(rate = sum(total) /  sum(population) * 10^6) %>%
  pull(rate)
r
```

## Visualize the data

A color coordinated graph of each state compared to the average gun murder rate in log scale.

```{r log-scale-graph}
murders %>% ggplot(aes(population/10^6, total, label = abb)) +
  geom_abline(intercept = log10(r), lty = 2, color = "darkgrey") +
  geom_point(aes(color=region), size = 3) +
  geom_text_repel() + 
  scale_x_log10() + 
  scale_y_log10() + 
  xlab("Population in millions (log scale)") +
  ylab("Total number of murders") +
  ggtitle("US Gun Murders in 2010") + 
  scale_color_discrete(name = "Region") + 
  theme_economist()
```

